How to measure the radial velocity of a star:\\The invariant-spectrum case

### 2022-02-15 paper/main.tex
2 more HOGG paragraphs

Because data are noisy, there are limits to the precision (and accuracy\footnote{HOGG}) of any measurement that you can make with those data.
The Cram\'er-Rao lower bound (CRLB; \citealt{crlb}) is a quantitative result; it says that the variance of any unbiased estimator $\widehat{\theta}$ of any quantity $\theta$ given noisy data $y$ has a variance greater than or equal to the inverse of the Fisher information $I_\theta$.
The Fisher Information is an expectation of the second derivative of the log-likelihood function with respect to the parameter $\theta$ (\citealt{fisher}); it can only be calculated in the context of a set of assumptions sufficient to deliver a specific likelihood function.
The CLRB is a frequentist bound, because the variance of the estimator, in this case, is the variance obtained in a hypothetical program of repeated experiments (holding everything but the noise draw fixed).
If we make an unbiased measurement $\widehat{\theta}$ of parameter $\theta$ given data $y$ (where this $y$ might be a list or set or vector of observations), that measurement $\widehat{\theta}$ will have expected variance $\sigma_\theta^2$, subject to an inequality:
\begin{align}
    \frac{1}{\sigma_\theta^2} &\leq I_\theta\label{eq:CRLB}\\
    I_\theta &\equiv -\E\left[\frac{\dd^2}{\dd\theta^2}\ln L\right]\label{eq:fisher}\\
    \ln L &\equiv \ln p(y\given\theta)~,
\end{align}
where the first line \eqref{eq:CRLB} expresses the CRLB,
the expectation $\E[\cdot]$ is taken over all possible data sets $y$,
and the likelihood function $L$ is the probability density for the data $y$ given the parameter $\theta$.
Note that since the Fisher information $I_\theta$ involves a second derivative of a dimensionless quantity (a logarithm) with respect to the parameter $\theta$, it has units of inverse-$\theta$-squared, which makes the CRLB consistent in a dimensional sense.

@@

HOGG: Quantitative argument using SNR and scalings.

@@

HOGG: Note the sensitivity to edge effects.... How did we deal with that?

>>>>

Because data are noisy, there are limits to the precision (and accuracy) of any measurement that you can make with those data.
The Cram\'er-Rao lower bound (CRLB; \citealt{crlb}) is a quantitative result; it says that the variance of any unbiased estimator $\widehat{\theta}$ of any quantity $\theta$ given noisy data $y$ has a variance greater than or equal to the inverse of the Fisher information $I_\theta$.
The Fisher Information is an expectation of the second derivative of the log-likelihood function with respect to the parameter $\theta$ (\citealt{fisher}); it can only be calculated in the context of a set of assumptions sufficient to deliver a specific likelihood function.
The CLRB is a frequentist bound, because the variance of the estimator, in this case, is the variance obtained in a hypothetical program of repeated experiments (holding everything but the noise draw fixed).
If we make an unbiased measurement $\widehat{\theta}$ of parameter $\theta$ given data $y$ (where this $y$ might be a list or set or vector of observations), that measurement $\widehat{\theta}$ will have expected variance $\sigma_\theta^2$, subject to an inequality:
\begin{align}
    \frac{1}{\sigma_\theta^2} &\leq I_\theta\label{eq:CRLB}\\
    I_\theta &\equiv -\E\left[\frac{\dd^2}{\dd\theta^2}\ln L\right]\label{eq:fisher}\\
    \ln L &\equiv \ln p(y\given\theta)~,
\end{align}
where the first line \eqref{eq:CRLB} expresses the CRLB,
the expectation $\E[\cdot]$ is taken over all possible data sets $y$,
and the likelihood function $L$ is the probability density for the data $y$ given the parameter $\theta$.
Note that since the Fisher information $I_\theta$ involves a second derivative of a dimensionless quantity (a logarithm) with respect to the parameter $\theta$, it has units of inverse-$\theta$-squared, which makes the CRLB consistent in a dimensional sense.

@@

There is a simple argument about the quality of empirical templates that goes as follows:
Imagine that you have $N$ epochs of observation of a star, all with the same spectrograph, and each of these observations is at similar signal-to-noise ratio.
If the star does not vary epoch-to-epoch, an empirical spectral template will accurately describe the spectrum of the star in this spectrograph (by assumption and construction), and it will have a signal-to-noise ratio that is on the order of $\sqrt{N}$ larger than the signal-to-noise of any individual epoch.
The use of the empirical template will add in quadrature to any individual-epoch Doppler-shift uncertainty a template-origin uncertainty that is smaller by a factor of $\sqrt{N}$.
Once $N$ is large enough---and because this additional noise adds only in quadrature---the empirical template ought to be as good (or better, since a certain kind of accuracy is guaranteed) as any theoretical spectral template for the determination of Doppler shifts.
In the toy data we use here (\secref{sec:data}), $N=64$.

@@

The empirical template can be subject to issues at its edges, since each spectral epoch reveals a slightly different stellar rest-frame spectral range in a spectrograph of fixed spectrograph-frame spectral range.
For this reason, when we made the spectral template, we zeroed out (or really set to unity) the empirical spectra in a logarithmic wavelength range that is within $40\,\kmps$ of either edge of the template spectral range.
Evidence of this censoring is visible in \figref{fig:empirical}.
